---
title: "Acropora cervicornis breakage model"
author: "Nick Jones"
date: "2024-05-02"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(plotrix)
library(tidyverse)
library(stringr)
library(moments)
library(vegan)
library(randomForest)
library(varImp)
library(pdp)
library(MESS)
library(glmmTMB)
library(DHARMa)
library(emmeans)
library(performance)
library(corrplot)
library(lattice)

```

```{r}

predictors <- read.csv("C:\\Users\\nj350\\OneDrive - Nova Southeastern University\\Documents\\Grants and Funding\\2023\\FY 24 DEP Restoration Hub\\Data\\Site Data\\All Predictors.csv", header = T)

breakage <- read.csv("C:\\Users\\nj350\\OneDrive - Nova Southeastern University\\Documents\\Grants and Funding\\2023\\FY 24 DEP Restoration Hub\\Data\\Outplant Data\\Data Summaries\\Array Breakage Prevalence.csv", header = T)

grid <- read.csv("C:\\Users\\nj350\\OneDrive - Nova Southeastern University\\Documents\\Grants and Funding\\2023\\FY 24 DEP Restoration Hub\\Data\\Outplant Data\\Data Summaries\\Grid Breakage Prevalence.csv", header = T)

```

### Data manipulation for analysis

```{r}

predictors$Site <- as.factor(predictors$Site)
breakage$Site <- as.factor(breakage$Site)
grid$Site <- as.factor(grid$Site)

breakage$Array <- as.factor(breakage$Array)
grid$Array <- as.factor(grid$Array)

breakage$Species <- as.factor(breakage$Species)
grid$Species <- as.factor(grid$Species)

breakage <- rbind(breakage, grid)

breakage.site <- left_join(predictors, breakage, by = c("Site", "Array"))
breakage.site$Array <- as.factor(breakage.site$Array)

#remove arrays with no acer, note G16 a Exp5 gets removed because it was an experimental site so didn't get included in this analysis
breakage.site <- na.omit(breakage.site)

###remove collinear variables - mean length and kurtosis

breakage.site <- breakage.site %>%
  select(c(-17,-19))

```

### Data analysis in 2 stage process - first random forest, then GLMM using most important variables
### Random Forest analysis - ACER breakage within year 1

```{r}

all.pred <- predictors %>%
  select(c(5, 10:28))

#check if correlation coefficient over 90 and remove one of the collinear variables
M <- cor(all.pred, method = "spearman") #spearman good for non-normal data with outliers
#remove LONDD (collinear with LatDD), mean.length (collinear with CV), kurtosis (collinear with skew)
corrplot(M, method = 'number')

```

```{r}

breakage.site$Method <- as.factor(breakage.site$Method)
breakage.site$Cluster <- as.factor(breakage.site$Cluster)

set.seed(1985)

forest.full <- randomForest(Breakage.Prev ~ n.x + Acer + Skew + CV + H + J + Sp.rich + 
                           Rugosity + Sed.max + Sed.mean + Relief.max + Relief.mean +
                           Distance.to.Shore + LatDD + Distance.to.Port + Depth + Cluster + Method,
                         data = breakage.site, importance = T,
                         replace = T, ntree = 500)

round(importance(forest.full),2)
plot(forest.full) #displays cross validation error rate against the number of trees to assess whether the forest is reliable
print(forest.full) #too complex - need to simplify, performed stepwise regression to find optimal starting model

set.seed(1985)
forest.1 <- randomForest(Breakage.Prev ~ n.x + J +  
                           Relief.max + LatDD,
                         data = breakage.site, importance = T,
                         replace = T, ntree = 500)

round(importance(forest.1),2)
plot(forest.1) #displays cross validation error rate against the number of trees to assess whether the forest is reliable
print(forest.1) #better but not accounting for much variation - 3.27%

#breakage predictions of the random forest at each array based upon model
p.rf <- predict(forest.1, newdata=breakage.site)
length(unique(p.rf))

#calculate the root mean squared error using the out of bag predictions- how far predicted values are from observed values in the regression analysis (how concentrated the data are around line of best fit)
#summarises how different the fitted predictions are with the actual data, quantiles and extremes are - rpp cntains each data point comparison
summary(r.rpp <- breakage.site$Breakage.Prev - p.rf) #mean is good, extremes off

rmse.rf <- sqrt(sum(r.rpp^2)/length(r.rpp)) #sqrt of the sum of squareds/length

summary(breakage.site$Breakage.Prev)
summary(p.rf) #model isn't great - high values, mean and median quite a way off

#Plot predicted vs actual values
ggplot(breakage.site, aes(x = Breakage.Prev, y = p.rf)) +
         geom_point() +
  geom_abline()+
         labs(x = "Actual", y = "Fitted")+
         scale_y_continuous(limits = c(0,50))+
         scale_x_continuous(limits = c(0,50)) #a little off but reasonably consistent, underestimates mortality at high values in particular

set.seed(2509)
forest.2 <- randomForest(Breakage.Prev ~ n.x + J +  
                           Relief.max + LatDD,
                         data = breakage.site, importance = T,
                         replace = T, ntree = 1000)

round(importance(forest.2),2)
print(forest.2)

p.rf2 <- predict(forest.2, newdata=breakage.site)
length(unique(p.rf))

summary(r.rpp <- breakage.site$Breakage.Prev - p.rf)

ggplot(breakage.site, aes(x = Breakage.Prev, y = p.rf2)) +
         geom_point(aes(colour = Site)) +
  geom_abline()+
         labs(x = "Actual", y = "Fitted")+
         scale_y_continuous(limits = c(0,50))+
         scale_x_continuous(limits = c(0,50)) #south commercial and north hub get a lot more breakage than you'd expect

```

### Extract Variable Importance

```{r}

#extracts variable importance
imp1 <- as.data.frame(round(importance(forest.1), 2)) #extracts variable importance
imp1$Var.Names <- row.names(imp1)
imp2 <- as.data.frame(round(importance(forest.2), 2))
imp2$Var.Names <- row.names(imp2)
importance <- full_join(imp1, imp2, by = "Var.Names") #joins var importance from both models
names(importance) <- c("Imp1", "Node.1", "Var.Names", "Imp2", "Node.2")
importance$Imp <- (importance$Imp1+importance$Imp2)/2
importance.order <- importance[order(-importance$Imp),] #orders variable importance 
cumsumbinning(importance.order$Imp, threshold = (sum(importance$Imp)*0.75), cutwhenpassed = T) #use first 6 variables
#looks at cutoffs at 75% total variable importance explained (e.g. vars with a 1 are in the top 75%, 2 and above are less important)
importance.order
mean(forest.1$mse) 
sum(importance$Imp)

#so R2 not high but we're looking to identify the most important variables

ggplot(importance, aes(x=Var.Names,  y=Imp)) +
  geom_segment(aes(x=reorder(Var.Names,Imp), xend=Var.Names, y=0, yend=Imp), color="grey") +
  geom_point(aes(color=Imp), size = 5) +
  labs(y = bquote("%IncMSE"))+
  theme_light() +
  coord_flip()+
  theme_classic()+
  theme(axis.text.y = element_text(size = 12, colour = "black"), 
        axis.title.y = element_blank(), 
        axis.text.x = element_text(size = 12, colour = "black"), 
        axis.title.x = element_text(size = 14),
        axis.line = element_line(linewidth = 1), 
        axis.ticks = element_blank(),legend.position="")+
  scale_color_continuous(type= "viridis")+
  scale_x_discrete(labels = c("Max relief height","Latitude",
                               "Wild coral abundance",  
                               "Evenness (J)"))

#ggsave(file = "ACER breakage var imp.png", width = 7, height = 4, dpi = 300) 
#J, n.x, Relief Max

```

### Re-run with the chosen variables that account for 75% of variation
```{r}

set.seed(25091985)

forest.3 <- randomForest(Breakage.Prev ~ n.x + J +  
                           Relief.max,
                         data = breakage.site, importance = T,
                         replace = T, ntree = 1000) 

round(importance(forest.3),2)
print(forest.3) #explains 3.17% variation

p.rf3 <- predict(forest.3, newdata=breakage.site)
summary(r.rpp <- breakage.site$Breakage.Prev - p.rf3) #extreme values off but mean is close and 1st and 3rd quantiles close

ggplot(breakage.site, aes(x = Breakage.Prev, y = p.rf3)) +
         geom_point(aes(colour = Site)) +
  geom_abline()+
         labs(x = "Actual", y = "Fitted")+
         scale_y_continuous(limits = c(0,50))+
         scale_x_continuous(limits = c(0,50)) #looks about the same in terms of accuracy

imp3 <- as.data.frame(round(importance(forest.3), 2))
imp3$Var.Names <- row.names(imp3)
names(imp3) <- c("Imp", "Node.3", "Var.Names")

ggplot(imp3, aes(x=Var.Names,  y=Imp)) +
  geom_segment(aes(x=reorder(Var.Names,Imp), xend=Var.Names, y=0, yend=Imp), color="grey") +
  geom_point(aes(color=Imp), size = 5) +
  labs(y = bquote("%IncMSE"))+
  theme_light() +
  coord_flip()+
  theme_classic()+
  theme(axis.text.y = element_text(size = 12, colour = "black"), 
        axis.title.y = element_blank(), 
        axis.text.x = element_text(size = 12, colour = "black"), 
        axis.title.x = element_text(size = 14),
        axis.line = element_line(linewidth = 1), 
        axis.ticks = element_blank(),legend.position="")+
  scale_color_continuous(type= "viridis") 
```

### Extract those main variables and look at relationship with breakage

```{r}

break.J <- forest.3 %>%  # the %>% operator is read as "and then"
  partial(pred.var = "J", plot.engine = "ggplot2") %>%
  plotPartial(smooth = TRUE, lwd = 2, ylab = expression(Breakage))
 # increasing breakage with evenness
break.J

break.n <- forest.3 %>%  # the %>% operator is read as "and then"
  partial(pred.var = "n.x", plot.engine = "ggplot2") %>%
  plotPartial(smooth = TRUE, lwd = 2, ylab = expression(Breakage))
 # increase with number of colonies
break.n 

break.Relief <- forest.3 %>%  # the %>% operator is read as "and then"
  partial(pred.var = "Relief.max", plot.engine = "ggplot2") %>%
  plotPartial(smooth = TRUE, lwd = 2, ylab = expression(Breakage))
 
break.Relief #high at low max relief, very low at moderate max relief

```

### GLMM of ACER breakage using the most important variables to fit model (inc interactions) and then make an equation that predicts survival prevalence
##### From random forest evidence of quadratic relationship between breakage prev and relief max in particular

```{r, results='hide'}

hist(breakage.site$Breakage.Prev)

breakage.site$R2 <- (breakage.site$Relief.max - mean(breakage.site$Relief.max))^2 #quadratic and centred CV
breakage.site$n2 <- (breakage.site$n.x - mean(breakage.site$n.x))^2 #quadratic and centred CV

#starting model - use interactions that think may be meaningful e.g., H:CV + H:Relief.mean - the spatial factors and outplanting methods are in as covariates
mod.1 <- glmmTMB(Breakage.Prev/100 ~ Relief.max + n.x + J + J:n.x + J:Relief.max + n.x:Relief.max + R2 + n2 +
                   (1|Site), 
                 weights = Initial.n, family = binomial, data = breakage.site)
summary(mod.1)

mod.2 <- glmmTMB(Breakage.Prev/100 ~ Relief.max + J + J + n.x + R2 +
                   (1|Site), 
                 weights = Initial.n, family = binomial, data = breakage.site)
summary(mod.2)

AIC(mod.1, mod.2)

mod.3 <- glmmTMB(Breakage.Prev/100 ~ Relief.max + n.x + J * n.x + R2 +
                   (1|Site), 
                 weights = Initial.n, family = binomial, data = breakage.site)
summary(mod.3)

mod.4 <- glmmTMB(Breakage.Prev/100 ~ Relief.max + n.x + J + n2 + J:Relief.max + R2 +
                   (1|Site), 
                 weights = Initial.n, family = binomial, data = breakage.site)
summary(mod.4)

mod.5 <- glmmTMB(Breakage.Prev/100 ~ n.x + n2 + J * n.x +
                   (1|Site), 
                 weights = Initial.n, family = binomial, data = breakage.site)

mod.6 <- glmmTMB(Breakage.Prev/100 ~ n.x + J + n2 + 
                   (1|Site), 
                 weights = Initial.n, family = binomial, data = breakage.site)

mod.7 <- glmmTMB(Breakage.Prev/100 ~ Relief.max + J * Relief.max + R2 +
                   (1|Site), 
                 weights = Initial.n, family = binomial, data = breakage.site)

mod.8 <- glmmTMB(Breakage.Prev/100 ~ Relief.max + J + R2 +
                   (1|Site), 
                 weights = Initial.n, family = binomial, data = breakage.site)

mod.9 <- glmmTMB(Breakage.Prev/100 ~ Relief.max + R2 +
                   (1|Site), 
                 weights = Initial.n, family = binomial, data = breakage.site)

mod.10 <- glmmTMB(Breakage.Prev/100 ~ Relief.max + n.x + n2 + R2 + n.x:Relief.max +
                   (1|Site), 
                 weights = Initial.n, family = binomial, data = breakage.site)

summary(mod.10)

mod.11 <- glmmTMB(Breakage.Prev/100 ~ Relief.max + n.x + n2 + R2 +
                   (1|Site), 
                 weights = Initial.n, family = binomial, data = breakage.site)

mod.12 <- glmmTMB(Breakage.Prev/100 ~ n.x + n2+
                   (1|Site), 
                 weights = Initial.n, family = binomial, data = breakage.site)

mod.13 <- glmmTMB(Breakage.Prev/100 ~ J +
                   (1|Site), 
                 weights = Initial.n, family = binomial, data = breakage.site)

AIC(mod.1, mod.2, mod.3, mod.4, mod.5, mod.6, mod.7, mod.8, mod.9, mod.10, mod.11, mod.12, mod.13)
#mod.3 best
summary(mod.3)

mod.3a <- glmmTMB(Breakage.Prev/100 ~ Relief.max + n.x + J * n.x + R2 +
                   (1|Site), 
                 weights = Initial.n, family = binomial, data = breakage.site)

AIC(mod.3, mod.3a)
#best model is Relief.max, colony abundance, Pielou J * colony abundance

summary(mod.3a)

```

### Fitted model - Breakage

```{r}

break.mod <- glmmTMB(Breakage.Prev/100 ~Relief.max + n.x + J * n.x + R2 +
                   (1|Site), 
                 weights = Initial.n, family = binomial, data = breakage.site)

sims1 <- simulateResiduals(break.mod, n = 1000)
plot(sims1) #model fit is good

e <- resid(break.mod)
plot(breakage.site$Relief.max, e)
plot(breakage.site$n.x, e)
plot(breakage.site$J, e)
plot(breakage.site$R2, e)  #the variable with very high relief, looks like an outlier so remove and refit

### remove outlier relief

breakage.site2 <- subset(breakage.site, Relief.max < 100)

break.mod2 <- glmmTMB(Breakage.Prev/100 ~ n.x + J * n.x + R2 +
                   (1|Site), 
                 weights = Initial.n, family = binomial, data = breakage.site2)
summary(break.mod2)
sims1 <- simulateResiduals(break.mod2, n = 1000)
plot(sims1) #model fit is good

e <- resid(break.mod2)
plot(breakage.site2$n.x, e)
plot(breakage.site2$J, e)
plot(breakage.site2$R2, e)

summary(break.mod2)

r2(break.mod2) ##conditional R2 = 0.13, marginal = 0.03 - so not very high

library(jtools) 

#extract model predictions and plot interaction
pred.mod <- predict(break.mod2, type = "response")
pred.df <- data.frame(breakage.site2, pred.mod)

ggplot(pred.df, aes(x = R2, y = Breakage.Prev))+
  geom_smooth()+
  geom_point(aes(size = pred.mod))

#https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html

#doesn't like the weights argument so fit this without to test interaction
break.mod.no.w <- glmmTMB(Breakage.Prev/100 ~ n.x + J * n.x + R2 +
                   (1|Site), family = binomial, data = pred.df)

mean(pred.df$J)
sd(pred.df$J)

interact.fig <- interact_plot(break.mod.no.w, pred = n.x, modx = J, data = pred.df,
              modx.values = c(0.6, 0.7, 0.8),
                            modx.labels = c("0.6 (-1 SD)", "0.7 (Mean)", "0.8 (+1 SD)"))+
    theme(legend.position = c(0.15,0.8),
        legend.text = element_text(size = 8, colour = "black"),
      legend.title = element_text(size = 10, colour = "black"),
        axis.title.y = element_text(size = 12, colour = "black"),
        axis.text.y = element_text(size = 10, hjust = 1, colour = "black"),  
        axis.text.x = element_text(size = 10, hjust = 1, colour = "black"),
        axis.title.x = element_text(size = 12, colour = "black"),
        axis.line = element_line(colour = "black", size = 0.5), 
        panel.border =  element_rect(size = 0.5, colour = "black", fill = NA),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        strip.background = element_blank(),
        strip.text = element_text(size = 12),
        panel.background = element_blank())+
    labs(y = bquote('Breakage Probability'), x = "Colony Abundance")
#increase in breakage with n.x, when J is above average becomes exponential increase 

interact.fig

#ggsave(interact.fig, file = "Breakage survival prevalence interaction.png", width = 7, height = 4, dpi = 300)

interact_plot(break.mod.no.w, pred = J, modx = n.x, data = pred.df)

```

### Forest plot of model
```{r}
plot_model(break.mod2, sort.est = T)+
  theme(legend.position = c(0.15,0.2),
        legend.text = element_text(size = 8, colour = "black"),
      legend.title = element_text(size = 10, colour = "black"),
        axis.title.y = element_text(size = 12, colour = "black"),
        axis.text.y = element_text(size = 10, hjust = 1, colour = "black"),  
        axis.text.x = element_text(size = 10, hjust = 1, colour = "black"),
        axis.title.x = element_text(size = 12, colour = "black"),
        axis.line = element_line(colour = "black", size = 0.5), 
        panel.border =  element_rect(size = 0.5, colour = "black", fill = NA),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        strip.background = element_blank(),
        strip.text = element_text(size = 12),
        panel.background = element_blank())

```