---
title: "Reef-building Outplants Model"
author: "Nick Jones"
date: "2024-05-08"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(plotrix)
library(tidyverse)
library(stringr)
library(moments)
library(vegan)
library(randomForest)
library(varImp)
library(pdp)
library(MESS)
library(glmmTMB)
library(DHARMa)
library(emmeans)
library(performance)
library(corrplot)
library(lattice)
library(sjPlot)
```

```{r}

predictors <- read.csv("C:\\Users\\nj350\\OneDrive - Nova Southeastern University\\Documents\\Grants and Funding\\2023\\FY 24 DEP Restoration Hub\\Data\\Site Data\\All Predictors.csv", header = T)

clusters <- read.csv("C:\\Users\\nj350\\OneDrive - Nova Southeastern University\\Documents\\Grants and Funding\\2023\\FY 24 DEP Restoration Hub\\Data\\Outplant Data\\Data Summaries\\Array Survival Prevalence.csv", header = T)

fwc <- read.csv("C:\\Users\\nj350\\OneDrive - Nova Southeastern University\\Documents\\Grants and Funding\\2023\\FY 24 DEP Restoration Hub\\Data\\Outplant Data\\Data Summaries\\FWC Bases.csv", header = T)

breakage <- read.csv("C:\\Users\\nj350\\OneDrive - Nova Southeastern University\\Documents\\Grants and Funding\\2023\\FY 24 DEP Restoration Hub\\Data\\Outplant Data\\Data Summaries\\Array Breakage Prevalence.csv", header = T)

grids <- read.csv("C:\\Users\\nj350\\OneDrive - Nova Southeastern University\\Documents\\Grants and Funding\\2023\\FY 24 DEP Restoration Hub\\Data\\Outplant Data\\Data Summaries\\Grid Array Survival Prevalence.csv", header = T)

```

### Data manipulation for analysis

```{r}

predictors$Site <- as.factor(predictors$Site)
clusters$Site <- as.factor(clusters$Site)
grids$Site <- as.factor(grids$Site)

clusters$Species <- as.factor(clusters$Species)
grids$Species <- as.factor(grids$Species)

fwc$Survival.Prev <- fwc$Survival.Prev*100

survival <- rbind(clusters, grids, fwc) %>%
  subset(Period != "0 Year")

survival <- left_join(predictors, survival, by = c("Site", "Array"))

###remove collinear variables

survival <- survival %>%
  select(c(-17,-19))

```

### Data analysis in 2 stage process - first random forest, then GLMM using most important variables
### Random Forest analysis - ACER survival 1 year

```{r}

all.pred <- predictors %>%
  select(c(5, 10:28))

#check if correlation coefficient over 90 and remove one of the collinear variables
M <- cor(all.pred, method = "spearman") #spearman good for non-normal data with outliers
#remove LONDD (collinear with LatDD), mean.length (collinear with CV), kurtosis (collinear with skew)
corrplot(M, method = 'number')

```


```{r}
reef.builders <- survival %>% subset(Species != "ACER" & Period == "1 Year")
reef.builders$Species <- droplevels(reef.builders$Species)
reef.builders$Method <- as.factor(reef.builders$Method)
reef.builders$Cluster <- as.factor(reef.builders$Cluster)

summary(reef.builders$Survival.Prev) #big range in 1 year survival

set.seed(1985)

forest.1 <- randomForest(Survival.Prev ~ Species + n.x + Skew + CV + H + J + Sp.rich + 
                           Rugosity + Sed.max + Sed.mean + Relief.max + Relief.mean +
                           Distance.to.Shore + LatDD + Distance.to.Port + Depth + Cluster + Method,
                         data = reef.builders, importance = T,
                         replace = T, ntree = 1000)

plot(forest.1) #displays cross validation error rate against the number of trees to assess whether the forest is reliable
print(forest.1) #R2 = 22.56

#survival predictions of the random forest at each array based upon model
p.rf <- predict(forest.1, newdata=reef.builders)
length(unique(p.rf))

#calculate the root mean squared error using the out of bag predictions- how far predicted values are from observed values in the regression analysis (how concentrated the data are around line of best fit)
#summarises how different the fitted predictions are with the actual data, quantiles and extremes are - rpp cntains each data point comparison
summary(r.rpp <- reef.builders$Survival.Prev - p.rf) #mean is good, extremes off, low survival is overestimated, high survival is underestimated

rmse.rf <- sqrt(sum(r.rpp^2)/length(r.rpp)) #sqrt of the sum of squareds/length = distance between actual and predicted values - 11.4%

summary(reef.builders$Survival.Prev)
summary(p.rf) #model is more conservative - lower values quite a way off, mean and median good

#Plot predicted vs actual values
ggplot(reef.builders, aes(x = Survival.Prev, y = p.rf)) +
         geom_point() +
  geom_abline()+
         labs(x = "Actual", y = "Fitted")+
         scale_y_continuous(limits = c(0,100))+
         scale_x_continuous(limits = c(0,100)) #actually looks pretty good, underestimates mortality at low values

set.seed(2509)
forest.2 <- randomForest(Survival.Prev ~ Species + n.x + Skew + CV + H + J + Sp.rich + 
                           Rugosity + Sed.max + Sed.mean + Relief.max + Relief.mean +
                           Distance.to.Shore + LatDD + Distance.to.Port + Depth + Cluster + Method,
                         data = reef.builders, importance = T,
                         replace = T, ntree = 1000)

round(importance(forest.2),2)
print(forest.2) #22.8

p.rf2 <- predict(forest.2, newdata=reef.builders)
length(unique(p.rf))

summary(r.rpp <- reef.builders$Survival.Prev - p.rf2)

ggplot(reef.builders, aes(x = Survival.Prev, y = p.rf2)) +
         geom_point(aes(colour = Site)) +
  geom_abline()+
         labs(x = "Actual", y = "Fitted")+
         scale_y_continuous(limits = c(0,100))+
         scale_x_continuous(limits = c(0,100)) #pretty good, apart from the very low survival of SBOU at exp5

```
### Extract Variable Importance

```{r}

#extracts variable importance
imp1 <- as.data.frame(round(importance(forest.1), 2)) #extracts variable importance
imp1$Var.Names <- row.names(imp1)
imp2 <- as.data.frame(round(importance(forest.2), 2))
imp2$Var.Names <- row.names(imp2)
importance <- full_join(imp1, imp2, by = "Var.Names") #joins var importance from both models
names(importance) <- c("Imp1", "Node.1", "Var.Names", "Imp2", "Node.2")
importance$Imp <- (importance$Imp1+importance$Imp2)/2
importance.order <- importance[order(-importance$Imp),] #orders variable importance 
cumsumbinning(importance.order$Imp, threshold = (sum(importance$Imp)*0.75), cutwhenpassed = T) #use first 9 variables
#looks at cutoffs at 75% total variable importance explained (e.g. vars with a 1 are in the top 75%, 2 and above are less important)
importance.order
mean(forest.1$mse) 
sum(importance$Imp)

#so R2 not high but we're looking to identify the most important variables

ggplot(importance, aes(x=Var.Names,  y=Imp)) +
  geom_segment(aes(x=reorder(Var.Names,Imp), xend=Var.Names, y=0, yend=Imp), color="grey") +
  geom_point(aes(color=Imp), size = 5) +
  labs(y = bquote("%IncMSE"))+
  theme_light() +
  coord_flip()+
  theme_classic()+
  theme(axis.text.y = element_text(size = 12, colour = "black"), 
        axis.title.y = element_blank(), 
        axis.text.x = element_text(size = 12, colour = "black"), 
        axis.title.x = element_text(size = 14),
        axis.line = element_line(linewidth = 1), 
        axis.ticks = element_blank(),legend.position="")+
  scale_color_continuous(type= "viridis") 
#species, latdd, coral abundance, rugosity, mean relief most important, then distance from shore/port, then J and CV

```
### Re-run with the chosen variables that account for 75% of variation
```{r}

set.seed(25091985)

forest.3 <- randomForest(Survival.Prev ~ Species + LatDD +  n.x + Rugosity +
                           Relief.mean + Distance.to.Shore + Distance.to.Port + J + CV,
                         data = reef.builders, importance = T,
                         replace = T, ntree = 1000) 

round(importance(forest.3),2)
print(forest.3) #22.7%
plot(forest.3)

imp3 <- as.data.frame(round(importance(forest.3), 2))
imp3$Var.Names <- row.names(imp3)
names(imp3) <- c("Imp", "Node", "Var.Names")
imp3 <- imp3[order(-imp3$Imp),] #orders variable importance 

cumsumbinning(imp3$Imp, threshold = (sum(imp3$Imp)*0.75), cutwhenpassed = T) #use first 6 variables

forest.4 <- randomForest(Survival.Prev ~ Species + LatDD +  n.x + Rugosity +
                           Relief.mean + Distance.to.Shore,
                         data = reef.builders, importance = T,
                         replace = T, ntree = 1000) 

round(importance(forest.4),2)
print(forest.4) #27.0%
plot(forest.4)

imp4 <- as.data.frame(round(importance(forest.4), 2))
imp4$Var.Names <- row.names(imp4)
names(imp4) <- c("Imp", "Node", "Var.Names")
imp4 <- imp4[order(-imp4$Imp),] #orders variable importance 

cumsumbinning(imp4$Imp, threshold = (sum(imp4$Imp)*0.75), cutwhenpassed = T) #use first 6 variables

p.rf4 <- predict(forest.4, newdata=reef.builders)
summary(r.rpp <- reef.builders$Survival.Prev - p.rf4) #extreme values off but mean is close and 1st and 3rd quantiles close

ggplot(reef.builders, aes(x = Survival.Prev, y = p.rf4)) +
         geom_point(aes(colour = Site)) +
  geom_abline()+
         labs(x = "Actual", y = "Fitted")+
         scale_y_continuous(limits = c(0,100))+
         scale_x_continuous(limits = c(0,100)) #looks about the same in terms of accuracy

ggplot(imp4, aes(x=Var.Names,  y=Imp)) +
  geom_segment(aes(x=reorder(Var.Names,Imp), xend=Var.Names, y=0, yend=Imp), color="grey") +
  geom_point(aes(color=Imp), size = 5) +
  labs(y = bquote("%IncMSE"))+
  theme_light() +
  coord_flip()+
  theme_classic()+
  theme(axis.text.y = element_text(size = 12, colour = "black"), 
        axis.title.y = element_blank(), 
        axis.text.x = element_text(size = 12, colour = "black"), 
        axis.title.x = element_text(size = 14),
        axis.line = element_line(linewidth = 1), 
        axis.ticks = element_blank(),legend.position="")+
  scale_color_continuous(type= "viridis") +
  scale_x_discrete(labels = c("Rugosity", 
                              "Mean relief height","Distance from shore",
                               "Wild coral abundance",  
                              "Latitude", "Species"))

#ggsave(file = "Boulder survival var imp.png", width = 7, height = 4, dpi = 300)

```
### Extract those main variables and look at relationship with survival

```{r}

reef.sp <- forest.4 %>%  # the %>% operator is read as "and then"
  partial(pred.var = "Species", plot.engine = "ggplot2") %>%
  plotPartial(smooth = TRUE, lwd = 2, ylab = expression(Survival))
 # large variation in survival by species
reef.sp

reef.lat <- forest.4 %>%  # the %>% operator is read as "and then"
  partial(pred.var = "LatDD", plot.engine = "ggplot2") %>%
  plotPartial(smooth = TRUE, lwd = 2, ylab = expression(Survival))
 # marked increase in survival above 26.00 deg lat
reef.lat 

reef.n <- forest.4 %>%  # the %>% operator is read as "and then"
  partial(pred.var = "n.x", plot.engine = "ggplot2") %>%
  plotPartial(smooth = TRUE, lwd = 2, ylab = expression(Survival))
 # linear decrease in survival with abundance once above 100
reef.n

reef.shore <- forest.4 %>%  # the %>% operator is read as "and then"
  partial(pred.var = "Distance.to.Shore", plot.engine = "ggplot2") %>%
  plotPartial(smooth = TRUE, lwd = 2, ylab = expression(Survival))
 # linear decline in survival with distance from shore
reef.shore

reef.relief <- forest.4 %>%  # the %>% operator is read as "and then"
  partial(pred.var = "Relief.mean", plot.engine = "ggplot2") %>%
  plotPartial(smooth = TRUE, lwd = 2, ylab = expression(Survival))
 # large increase in survival once relief above 10cm
reef.relief

reef.rug <- forest.4 %>%  # the %>% operator is read as "and then"
  partial(pred.var = "Rugosity", plot.engine = "ggplot2") %>%
  plotPartial(smooth = TRUE, lwd = 2, ylab = expression(Survival))
 # major increase around 0.9 - basically around mean rugosity, relatively high rugosity or flat areas low survival, fit quadratic
reef.rug
summary(predictors$Rugosity)

```
### GLMM of reef.building species survival using the most important variables to fit model (inc interactions) and then make an equation that predicts survival prevalence
##### From random forest evidence of quadratic relationship between survival prev and rugosity

```{r}

hist(reef.builders$Survival.Prev)

reef.builders$R2 <- (reef.builders$Rugosity - mean(reef.builders$Rugosity))^2 #quadratic and centred CV

#we know there's loads of species variance but not the main focus therefore fitted as a random effect
#full model
mod.1 <- glmmTMB(Survival.Prev/100 ~ LatDD + Distance.to.Shore + Rugosity + 
                   n.x * Relief.mean * R2 + 
                   (1|Site) + (1|Species), 
                 weights = Initial.n, family = binomial, data = reef.builders)
summary(mod.1) # 3 way interaction was too complex - therefore edited
r2(mod.1)

```
### Model selection

```{r, results='hide'}

mod.2 <- glmmTMB(Survival.Prev/100 ~ LatDD + Distance.to.Shore + Rugosity + 
                   n.x + Relief.mean + R2 +
                   n.x : Relief.mean + Relief.mean:R2 +  
                   (1|Site) + (1|Species), 
                 weights = Initial.n, family = binomial, data = reef.builders)
summary(mod.2)

mod.3 <- glmmTMB(Survival.Prev/100 ~ LatDD + Distance.to.Shore + Rugosity + 
                   n.x + Relief.mean + R2 +
                   n.x : Relief.mean +   
                   (1|Site) + (1|Species), 
                 weights = Initial.n, family = binomial, data = reef.builders)
summary(mod.3)

mod.4 <- glmmTMB(Survival.Prev/100 ~ Distance.to.Shore + Rugosity + 
                   n.x + Relief.mean + R2 +
                   n.x : Relief.mean +   
                   (1|Site) + (1|Species), 
                 weights = Initial.n, family = binomial, data = reef.builders)
summary(mod.4)

mod.5 <- glmmTMB(Survival.Prev/100 ~ Rugosity + 
                   n.x + Relief.mean + R2 +
                   n.x : Relief.mean +   
                   (1|Site) + (1|Species), 
                 weights = Initial.n, family = binomial, data = reef.builders)
summary(mod.5) #can't really simplify further through backwards selection - could try best approach?

AIC(mod.1, mod.2, mod.3, mod.4, mod.5) #basically equivalent models - use mod.5

```
### Model Validation

```{r}

mod.fit <- glmmTMB(Survival.Prev/100 ~ Rugosity + 
                   n.x + Relief.mean + R2 +
                   n.x : Relief.mean +   
                   (1|Site) + (1|Species), 
                 weights = Initial.n, family = binomial, data = reef.builders)
summary(mod.fit) #negative with rugosity, R2, mean relief, n.x, interaction between n.x and relief.mean
r2(mod.fit) #0.355, 0.063

sims <- simulateResiduals(mod.fit, n = 1000)
plot(sims)#not brilliant, uniform, linearity fine, some heterogeneity in residuals
check_collinearity(mod.fit) # interaction terms only collinear bit. keep all as per recommendations in Zuur et al 2009 and online stat.exchange

testDispersion(sims) #dispersion fine
testOutliers(sims)# no outliers

e <- resid(mod.fit)
plot(reef.builders$Relief.mean, e)
plot(reef.builders$n.x, e)
plot(reef.builders$Rugosity, e)
plot(reef.builders$R2, e) #spreads pretty high but dispersion reasonably consistent apart from some high relief values - could remove them

```

```{r}

#extract model predictions and plot interaction
pred.mod <- predict(mod.fit, type = "response")
pred.df <- data.frame(reef.builders, pred.mod)

ggplot(pred.df, aes(y = pred.mod, x = R2))+
  geom_smooth()

library(jtools) 
library(interactions)
#https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html

#doesn't like the weights argument so fit this without to test interaction, look at simple model of interactions
mod.reef <- glmmTMB(Survival.Prev/100 ~ Rugosity + 
                   n.x + Relief.mean + R2 +
                   n.x : Relief.mean +   
                   (1|Site) + (1|Species), family = binomial, data = pred.df)

interact.fig <- interactions::interact_plot(mod.reef, 
                            pred = n.x, modx = Relief.mean, 
                            data = pred.df, legend.main = "Mean Relief Height", modx.values = c(8, 15.5, 23),
                            modx.labels = c("8 cm (-1 SD)", "15.5 cm (Mean)", "23 cm (+1 SD)"))+
  theme(legend.position = c(0.15,0.2),
        legend.text = element_text(size = 8, colour = "black"),
      legend.title = element_text(size = 10, colour = "black"),
        axis.title.y = element_text(size = 12, colour = "black"),
        axis.text.y = element_text(size = 10, hjust = 1, colour = "black"),  
        axis.text.x = element_text(size = 10, hjust = 1, colour = "black"),
        axis.title.x = element_text(size = 12, colour = "black"),
        axis.line = element_line(colour = "black", size = 0.5), 
        panel.border =  element_rect(size = 0.5, colour = "black", fill = NA),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        strip.background = element_blank(),
        strip.text = element_text(size = 12),
        panel.background = element_blank())+
    labs(y = bquote('Survival Probability'), x = "Colony Abundance")

interact.fig

#ggsave(interact.fig, file = "Reef builder survival prevalence.png", width = 7, height = 4, dpi = 300)

mean(pred.df$Relief.mean)
sd(pred.df$Relief.mean)

interactions::interact_plot(mod.reef, 
                            pred = Relief.mean, modx = n.x, 
                            data = pred.df)#survival prevalence declines with n.x at low relief and medium relief, but increases with n.x when relief about average

```

### Forest plot of model
```{r}
plot_model(mod.fit, sort.est = T)+
  theme(legend.position = c(0.15,0.2),
        legend.text = element_text(size = 8, colour = "black"),
      legend.title = element_text(size = 10, colour = "black"),
        axis.title.y = element_text(size = 12, colour = "black"),
        axis.text.y = element_text(size = 10, hjust = 1, colour = "black"),  
        axis.text.x = element_text(size = 10, hjust = 1, colour = "black"),
        axis.title.x = element_text(size = 12, colour = "black"),
        axis.line = element_line(colour = "black", size = 0.5), 
        panel.border =  element_rect(size = 0.5, colour = "black", fill = NA),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        strip.background = element_blank(),
        strip.text = element_text(size = 12),
        panel.background = element_blank())

```
